Enforced Invariant Confluence

### 2020-08-08 doc/the_vldb_journal_paper/rebuttal.tex

We describe this in Section 6.2, for example, when we say that if ``$t \notin
T_i$, then $p_i$ initiates a round of global coordination to execute $t$.
During global coordination, every server temporarily stops processing
transactions.''

Your question also reinforces the conclusion of Figure 10. If coordination is
frequent, then segmented \invariantconfluence{} becomes slow, even slower than
linearizable replication. Segmented \invariantconfluent{} works well only when
coordination is infrequent.

\paragraph{Reviewer 1: D13}
\begin{feedback}
  In 8.1 again, at the end, more discussion about the tradeoff of
  fault-tolerance would be interesting.
\end{feedback}
See Reviewer 4: 5, 7.

\paragraph{Reviewer 1: Typos}
\begin{feedback}
  end of first column of page 3 ``that it execute a transaction'' --> ``that it
  executes a transaction''

  section 5, in example 3, ``Our start state state'' --> two times state

  section 8.2 in example 12: ``and a optional'' --> ``and an optional''
\end{feedback}
Fixed.

@@

\paragraph{Reviewer 3: Section 7}
\begin{feedback}
  The authors added Section 7 to highlight a dual ``operation-based'' approach.
  This is not a significant improvement to the original contribution.
\end{feedback}
We disagree. The dual operation-based definitions of \invariantconfluence{} are
both theoretically interesting and practically important. Theoretically, as
with work on CRDTs, having dual models of the same concept can be helpful in
writing proofs and building intuition. Some properties are more naturally
expressed in the state-based model, while others are more natural in the
operation-based model. The operation-based model also has practical importance.
Many CRDTs are easier to understand as state-based objects, but are more
efficiently implemented as operation-based objects. By introducing
operation-based definitions of \invariantconfluence{}, we allow system
designers to choose whichever form of CRDT is more efficient. As we describe in
the paper, for example, the operation-based definitions let us replace a
PN-counter with a more efficient integer.

\paragraph{Reviewer 3: Reviewer 3: Other Comments}
\begin{feedback}
  Figure 5 should appear before Figure 6.
\end{feedback}
Fixed!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer 4}
\paragraph{Reviewer 4: Overview}
\begin{feedback}
  The paper is very well written, and it makes substantial improvement over the
  VLDB 2019 conference version.
\end{feedback}
Thanks! We appreciate it!

\paragraph{Reviewer 4: 1}
\begin{feedback}
  In the introduction the concept of ``invariant'' is missing some context. Where
  does this invariant come from? It looks like this invariant is a given
  invariant for the program executing on the replicated object, and it is not
  necessarily a tight invariant.  Clarifying this in advance would be useful,
  as I realized this only in the beginning of Section 2, I read this:
  ``Informally, a replicated object is invariant confluent *with respect to an
  invariant* if every replica of the object is guaranteed to satisfy the
  invariant despite the possibility of different replicas being concurrently
  modified or merged together.'' Mentioning the system model explicitly in the
  introduction also would help: ``a distributed object and accompanying
  invariant is replicated across a set of servers.''
\end{feedback}
We re-worded the introduction with your suggestion and also added a short
example of \invariantconfluence{} to the introduction.

@@

\paragraph{Reviewer 4: 2}
\begin{feedback}
  Invariant confluence is about safety, however, could there be a liveness part
  of the problem worth investigating/discussing for replicated object
  management? Would it be possible that forgoing coordination across replicas
  could jeopardize/violate some liveness properties?
\end{feedback}
Forgoing coordination across replicas in favor of \invariantconfluence{}
actually \emph{increases} liveness, rather than jeopardizing or violating it.
Any alive replica of an \invariantconfluent{} object can service client
requests and execute transactions. This is the strongest form of liveness we
can hope to achieve. In fact, if we aim for strongly consistent replication,
then the CAP theorem tells us that total availability (the kind that
\invariantconfluence{} achieves) is impossible~\cite{brewer2012cap,
gilbert2002brewer}.

Note, however, that a segmented \invariantconfluent{} object does suffer from
liveness problems, as we discuss in Section 6.4. Specifically, transactions
that require global coordination are live only as long as every replica is
live. We discuss this more in Reviewer 4: 5 below.

\paragraph{Reviewer 4: 3}
\begin{feedback}
  Mergeable Replicated Data Types paper (OOPSLA'19) is relevant work. The paper
  does not cite MRDT understandably because it is very recent work, however
  including a discussion of this in the related work would be helpful.
\end{feedback}
Thanks for pointing out this paper. We hadn't heard of it before, and it is
definitely relevant to our work on \invariantconfluence{}. We added a
discussion of MRDTs to our related works section.

### 2020-08-08 doc/the_vldb_journal_paper/sections/introduction.tex
More work on revisions.
Only two left!

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes...\ sometimes. Bailis et
al.\ introduced the notion of \emph{\invariantconfluence{}} as a necessary and
sufficient condition for when invariants can be maintained over replicated data
without the need for any coordination~\cite{bailis2014coordination}. If an
object is \invariantconfluent{} with respect to an invariant, we can replicate
it with the performance benefits of weak consistency and (some of) the
correctness benefits of strong consistency.

>>>>

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes...\ sometimes. Bailis et
al.\ introduced the notion of \emph{\invariantconfluence{}} as a necessary and
sufficient condition for when invariants can be maintained over replicated data
without the need for any coordination~\cite{bailis2014coordination}. If an
object is \invariantconfluent{} with respect to an \markrevisions{accompanying}
invariant, we can replicate the \markrevisions{object and invariant} with the
performance benefits of weak consistency and (some of) the correctness benefits
of strong consistency.
\begin{revisions}
  For example, a replicated bank account balance (represented as an integer) is
  \invariantconfluent{} with respect the invariant of a non-negative balance,
  so long as we only allow deposits.
\end{revisions}

### 2020-08-08 doc/the_vldb_journal_paper/sections/related_work.tex
Added tech report sections of paper.
More work on revisions.
Only two left!

@@

\begin{techreport}

  \textbf{Program Analysis in Database Systems.} Using program analysis to
  improve the performance of database systems is not new. For example, it has
  been used to improve the performance of database-backed web
  applications~\cite{cheung2014using, wu2016transaction,
  ramachandra2012program} and used to improve the performance of optimistic
  concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
  on \invariantconfluence{} continues the theme of using program analysis to
  reap the performance benefits gained from understanding program semantics.
\end{techreport}

>>>>

\begin{revisions}
  \textbf{Mergeable Replicated Data Types.}
  Mergeable Replicated Data Types (MRDTs)~\cite{kaki2019mergeable}, like CRDTs,
  are eventually consistent replicated objects. CRDTs get their eventual
  consistency by merging two CRDT replicas using a commutative, associative,
  and idempotent merge function. MRDTs instead merge two divergent replicas
  along with their least common ancestor. MRDTs rely on an abstraction function
  $\alpha$ and concretization function $\gamma$ to map abstract data types to
  the domain of relations. Given these two functions, one can derive an
  appropriate merge function.

  Like CRDTs, MRDTs are eventually consistent but may not preserve invariants,
  while \invariantconfluent{} objects preserve invariants but may not be
  eventually consistent. An interesting direction for future work would be to
  generalize \invariantconfluence{}'s use of binary merge operators to use the
  ternary merge operators used by MRDTs.
\end{revisions}

@@

\textbf{Program Analysis in Database Systems.} Using program analysis to
improve the performance of database systems is not new. For example, it has
been used to improve the performance of database-backed web
applications~\cite{cheung2014using, wu2016transaction,
ramachandra2012program} and used to improve the performance of optimistic
concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
on \invariantconfluence{} continues the theme of using program analysis to
reap the performance benefits gained from understanding program semantics.
